\begin{quote}
	And so, O Savior, give me at least some sign that I may know my path has not veered into madness, so I may, with that knowledge, walk the most difficult road, walk as long as need be and no longer feel weariness.

	What sign do you want and what knowledge? asked an elder standing by the Empty Tomb. Do you not know that any journey harbors danger within itself? Any journey—and if you do not acknowledge this, then why move? So you say faith is not enough for you and you want knowledge, too. But knowledge does not involve spiritual effort; knowledge is obvious. Faith assumes effort. Knowledge is repose and faith is motion.

	But were the venerable not aspiring for the harmony of repose? asked Arseny.

	They took the route of faith, answered the elder. And their faith was so strong it turned into knowledge.

	I want only to know the general direction of the journey, said Arseny. The part that concerns me and Ustina.

	But is not Christ a general direction? asked the elder. What other kind of direction do you seek? And how do you even understand the journey anyway? As the vast expanses you left behind? You made it to Jerusalem with your questions, though you could have asked them from the Kirillov Monastery. I am not saying wandering is useless: there is a point to it. Do not become like your beloved Alexander who had a journey but had no goal. And do not be enamored of excessive horizontal motion.

	Then what should I be enamored of? asked Arseny.

	Vertical motion, answered the elder, pointing above.

	In the center of the church’s cupola there gaped a round, black opening reserved for the sky and stars. Stars were visible but they were fading from sight. Arseny understood day was breaking.
	\attrib{\textit{Laurus}, Eugene Vodolazkin 2016}
\end{quote}