\begin{quote}
  The collapse [of the Ptolemaic Model] was not without collateral damage to the medieval mind. Along with this downgrading of the earth and its assumptive place in the cosmic hierarchy came the logical corollary that humankind was not as central as historically presumed. If this were so, then perhaps humankind's teleology and destiny were also in need of a downgrading. Life was just life, death just death, and perhaps the universe no longer deserved being "treated as a great allegory whose esssential secret was its religious meaning."
  \attrib{\textit{Theosis} 28, Michael Paul Gama 2017}
\end{quote}

\begin{quote}
  If there is any living theologican who might credibly serve as an icon of the truly modern theologian, it is likely Thomas Oden, who describes himself as a classical "Movement Theologican." His activism afforded him a front row seat on nearly innumerable bandwagons, including the early ecumenical movement, the civil rights struggles, the labor movement, engagement with the ACLU, forays into Gestalt therapy, and more. Many more.
  \attrib{\textit{Theosis} 31, Michael Paul Gama 2017}
\end{quote}

\begin{quote}
  Science replaced religion as preeminent intellectual authority as definer, judge, and guardian of the cultural world view. Human reason and empirical observation replaced theological doctrine and scriptural revelation as the principal means for comprehending the universe.
  \attrib{\textit{Passion} 270-271, Tarnas via \textit{Theosis} 32}
\end{quote}

\begin{quote}
  Throughout the modern era, intellectuals in many disciplines have turned to the reasoning subject rather than divine revelation as the starting point for knowledge and reflection Even modern theologians felt constrained to build on the foundation of rationalistic philosophy. They, too, accepted the primacy of reason advocated by Descartes... In the end, modern theologians ended up following Descartes's lead rather than trying to swim against the surge generated by the Age of Reason.
  \attrib{\textit{Primer} 65-66, Grenz via \textit{Theosis} 33}
\end{quote}

\begin{quote}
  "[The Puritans] had held that assurance is rare, late and the fruit of struggle in the experience of believers, the evangelicals believed it to be general, normally given at conversion and the result of simple acceptance of the gift of God"
  \attrib{\textit{Advent} 347, Haykin and Stewart via \textit{Theosis} 34}
\end{quote}

\begin{quote}
  Western Christianity went to sleep in a modern world governed by the gods of reason and observation. It is awakening to a postmodern world open to revelation and hungry for experience. Indeed, one of the last places postmoderns expect to be 'spiritual' is the church.
  \attrib{\textit{Post-Modern Pilgrims} 28, Sweet via \textit{Theosis} 86}
\end{quote}

\begin{quote}
  "[The Puritans] had held that assurance is rare, late and the fruit of struggle in the experience of believers, the evangelicals believed it to be general, normally given at conversion and the result of simple acceptance of the gift of God"
  \attrib{\textit{Advent} 347, Haykin and Stewart via \textit{Theosis} 34}
\end{quote}
